Suppl. Figure 3. & 4. Mirobiome-KEGG correlations.

### 2022-02-10 reports/rmarkdown/01_taxonomy_improved.Rmd

#Load libraries and functions

```{r, echo = FALSE}
library(tidyverse)
library(phyloseq)
library(Biostrings)
```

```{r function}
######################## FASTA Function ########################
write.fasta <- function(df, file.out, open = "w"){
  outfile <- file(description = file.out, open = "w")
  write.oneseq <- function(sequence, name){
    writeLines(paste(">", name, sep = ""), outfile)
    writeLines(sequence, outfile)
    return(name)
  }
  map2_chr(df$Sequence, df$Seq_ID, ~write.oneseq(.x, .y))
  close(outfile)
}

@@

# Loading data and metadata

```{r Get ASV and tax tables, message=FALSE}
paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region

seqtabs <- map(ps_list, ~data.frame(.x@otu_table@.Data))
taxtabs <- map(ps_list, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
sample_info <- map(ps_list, ~as_tibble(sample_data(.x), rownames = 'SampleID')) 
```

@@

```{r SeqID dict, include=FALSE, results='hide'}
############### Create dictonary for Sequence id ####
Seq_ID_key <-seqtab.taxa %>% 
  bind_rows(.id = "Sequencing_run") %>%
  # Get only uniqe sequences by grouping on all taxonomy
  group_by(Sequence,Kingdom,Phylum,Class,Order,Family,Genus,Species) %>%
  #summarize(n()) %>% filter(`n()`>1) %>% .$`n()` %>% sum(.)
  summarise(Reads_per_taxa = sum(Reads),
            Sequencing_run = paste(unique(Sequencing_run), collapse="; "),
            .groups = "drop") %>%
  bind_cols(Seq_ID = c(paste0("Seq_",sprintf("%05.0f", 1:nrow(.))))) %>%
  {. ->> seqtab.taxa.all } %>% 
  select(Seq_ID, Sequence)

@@

# BLAST
## Download databases

```{r Download data base}
db_dir <- "../../data/database/"
if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }

>>>>

# Load libraries, data and metadata

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
library(tidyverse)
library(phyloseq)
library(Biostrings)
})  })  })

#############
# LODA DATA #
#############
paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region

seqtabs <- map(ps_list, ~data.frame(.x@otu_table@.Data))
taxtabs <- map(ps_list, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
sample_info <- map(ps_list, ~as_tibble(sample_data(.x), rownames = 'SampleID'))

######################
# DOWNLOAD DATABASES #
######################
db_dir <- "../../data/database/"
if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }

```{r function}
################## 
# FASTA FUNCTION #
##################
write.fasta <- function(df, file.out, open = "w"){
  outfile <- file(description = file.out, open = "w")
  write.oneseq <- function(sequence, name){
    writeLines(paste(">", name, sep = ""), outfile)
    writeLines(sequence, outfile)
    return(name)
  }
  map2_chr(df$Sequence, df$Seq_ID, ~write.oneseq(.x, .y))
  close(outfile)
}

@@

@@

```{r SeqID dict, include=FALSE, results='hide'}
########################
# CREATE SEQUENCE ID's #
########################
Seq_ID_key <-seqtab.taxa %>% 
  bind_rows(.id = "Sequencing_run") %>%
  # Get only uniqe sequences by grouping on all taxonomy
  group_by(Sequence,Kingdom,Phylum,Class,Order,Family,Genus,Species) %>%
  #summarize(n()) %>% filter(`n()`>1) %>% .$`n()` %>% sum(.)
  summarise(Reads_per_taxa = sum(Reads),
            Sequencing_run = paste(unique(Sequencing_run), collapse="; "),
            .groups = "drop") %>%
  bind_cols(Seq_ID = c(paste0("Seq_",sprintf("%05.0f", 1:nrow(.))))) %>%
  {. ->> seqtab.taxa.all } %>% 
  select(Seq_ID, Sequence)

@@

# BLAST

### 2022-02-10 reports/rmarkdown/03_integrated_datasets.Rmd

>>>>

```{r eval=FALSE, include=FALSE}
groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
names(groupings) <- names(datasets_all_samples)[-1]
all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))

groupings_all_samples <- lapply(groupings,function(x){
  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
  df[rownames(x),1] <- as.character(x[,1])
  return(df)
})
names(groupings_all_samples) <- names(groupings)

df <- as.data.frame(groupings_all_samples)
colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
metadata <- cbind(metadata,df[rownames(metadata),])
```

### 2022-02-10 reports/rmarkdown/manuscript/Figure1.Rmd

---
title: "Figure 1.Luminal Study Groups"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

@@

```{r Figure 1., echo=FALSE, fig.asp=1, fig.height=7, fig.pos="t", fig.width=6, message=FALSE, warning=FALSE}
figlabels <- letters

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                5,6,7,7,8,8,
                5,4,7,7,8,8,
                5,4,7,7,8,8) , ncol= 6,byrow = T),
                               widths = c(1.87,1.42,.7,.8,.8,.7,.8),
                               heights = c(1.6,.65,1.5,.95,.9,.2))
### A
#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0,3.3,0,1))
fig1a <- magick::image_read("../../../supplementary_files/Scematic figure.pdf", density = 900 )
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

@@

  par(mar=c(1,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

@@

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T)

@@

###########
# LEGENDS #
###########
par(mar=c(1,0,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, y = 1.04, 
       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,3.3,1.5,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,1.5,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

@@

par(mar=c(0,0,1,1))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10, layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "Louvain", line = -2, cex.main = 1)

@@

### UMAP ###
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
U <- uwot::umap(adj,n_neighbors = 30)
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "UMAP", line = -2, cex.main = 1)

>>>>

---
title: "Figure 1.Luminal Study Groups"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(
  fig.path="./Figures/",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
```

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

@@

```{r Figure 1, echo=FALSE, fig.pos="t", fig.width=6, fig.height=5, fig.asp=1, message=FALSE, warning=FALSE}
figlabels <- letters
# fig.height=7, fig.asp=1,

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                5,6,7,7,8,8,
                5,4,7,7,8,8,
                5,4,7,7,8,8) , ncol= 6,byrow = T),
                               widths = c(1.87,1.42,.7,.8,.8,.7,.8),
                               heights = c(1.3,.55,1.33,.65,.9,0)) 
### A
#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0,3.3,0,1)) #b,l,t,r
fig1a <- magick::image_read("../../../supplementary_files/Scematic figure.pdf", density = 900 )
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

@@

  par(mar=c(1.5,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

@@

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T)

@@

###########
# LEGENDS #
###########
par(mar=c(0,0,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, y = 1.04, 
       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other"),
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,3.3,0,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,0,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

@@

par(mar=c(0,0,0,1)) #b,l,t,r
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10, layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "Louvain", line = -1, cex.main = 1)

@@

### UMAP ###
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
U <- uwot::umap(adj,n_neighbors = 30)
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "UMAP", line = -1, cex.main = 1)

### 2022-02-10 reports/rmarkdown/manuscript/Suppl.Figure3-4.Rmd

---
title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(openxlsx)
  library(fgsea)
  library(rafalib)
  library(Rcpp)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

```{r, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
myletter <- letters
#reprod. <- FALSE
reprod. <- TRUE

@@

  par(mfrow=c(1,1),mar=c(12,0,2,4)) #b,l,t,r
  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)

@@

**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown.Only bacterium and pathways with at least 10 significant NES scores were included in the heatmap.

@@

#Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = rev(sub("HALLMARK_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

#Combined representation of the two datasets
layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
par(mar=c(6,30,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]

>>>>

---
title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(
  fig.path="./Figures/",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(openxlsx)
  library(fgsea)
  library(rafalib)
  library(Rcpp)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
myletter <- letters
#reprod. <- FALSE
reprod. <- TRUE

@@

  ### Suppl. Figure 3-4
  #######################################
  # FUNCTIONAL ASSOCIATION BACT AND TRX #
  #######################################
  par(mfrow=c(1,1),mar=c(12,0,2,4)) #b,l,t,r
  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)

@@

**Suppl. Figure 3-4.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown.Only bacterium and pathways with at least 10 significant NES scores were included in the heatmap.

@@

# Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = rev(sub("HALLMARK_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

# Combined representation of the two datasets
layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
par(mar=c(6,30,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
